---
title: Add title here
author: Peter Carbonetto
date: June 18, 2020
site: workflowr::wflow_site
output: workflowr::wflow_html
---

*Add text here giving overview of analysis.*

```{r knitr, echo=FALSE}
knitr::opts_chunk$set(comment = "#",results = "hold",collapse = TRUE,
                      fig.align = "center",dpi = 140)
```

Analysis settings
-----------------

Analysis settings: number of simulations (ns), number of samples in
training set (n), number of simulated variables (p), number of
variables with a nonzero effect on the continuous outcome (p1),
proportion of variance in the outcome explained by the variables
(pve).

```{r params}
ns  <- 10
n   <- 500
p   <- 1000
p1  <- 200
pve <- 0.95
```

*Briefly explain here how these packages are being used.*

```{load-pkgs}
library(glmnet)
library(mr.ash.alpha)
library(ggplot2)
library(cowplot)
source("../code/functions_for_mr_ash_vs_lasso.R")
```

*Add text here describing how the simulations are implemented.*

```{r run-sims}

```
