---
title: "mr-ash demo"
author: "Peter Carbonetto"
date: February 20, 2020
site: workflowr::wflow_site
output: workflowr::wflow_html
---

An illustration of the "mr-ash" co-ordinate ascent algorithm applied
to a small, simulated data set.

```{r knitr, echo=FALSE}
knitr::opts_chunk$set(comment = "#",results = "hold",collapse = TRUE,
                      fig.align = "center")
```

## Script parameters

These are the data simulation settings.

```{r datasim-params}
n  <- 500
p  <- 1000
sd <- c(0,    1,    2)
w  <- c(0.98, 0.01, 0.01)
```

This specifies the variances for the mixture-of-normals prior on the regression
coefficients.

```{r set-s0}
s0 <- c(0.01,0.5,1)^2
```

## Load functions

This R code provides a very simple implementation of the mr-ash
algorithm.

```{r load-functions}
source("../code/misc.R")
source("../code/mr_ash.R")
```

## Simulate data

```{r simulate-data}
set.seed(1)
X    <- matrix(rnorm(n*p),n,p)
k    <- sample(length(w),p,replace = TRUE,prob = w)
beta <- sd[k] * rnorm(p)
y    <- drop(X %*% beta + rnorm(n))
```

## Fit model

These are the initial estimates of residual variance (`s`), mixture
weights (`w0`) and posterior mean estimates of the regression
coefficients (`b`).

```{r init-model}
b  <- rep(0,p)
s  <- 1
w0 <- c(0.5,0.25,0.25)
```

Fit the model by running 20 EM updates.

```{r fit-model}
fit <- mr_ash(X,y,s,s0,w0,b,20)
```

## Review model fit

Compare the mr-ash estimates against the values used to simulate the
data.

```{r plot-coefs, fig.height=4.5, fig.width=4}
plot(beta,fit$b,pch = 20,col = "black")
abline(a = 0,b = 1,col = "skyblue",lty = "dotted",xlab = "true",
       ylab = "estimated")
```

Plot the improvement in the model fit over time, as measured by the
ELBO.

```{r plot-progress, fig.height=4.5, fig.width=4}
plot(max(fit$elbo) - fit$elbo + 1e-5,type = "l",col = "dodgerblue",
     lwd = 2,log = "y",xlab = "iteration",ylab = "distance to \"best\" ELBO")
```
